\pkg{ggenealogy}: An \proglang{R} Package for Visualizing Genealogical Data

### 2017-04-15 article.tex
Ready for resubmission

Additionally, the \pkg{ggenealogy} package allows for users to plot genealogical data in multiple ways. Users can obtain and plot the shortest path between two nodes of interest constrained on a variable of interest (see Section \ref{psp}), superimpose a shortest path between two nodes of interest with the entire genealogical structure constrained on a variable of interest (see Section \ref{sspt}), plot the ancestors and descendants of a node of interest constrained on generation count (see Section \ref{remedy}), and plot distance matrices based on a variable of interest (see Section \ref{pdm}).

Most of these plotting tools were developed by building upon the \pkg{ggplot2} package (\citealt{ggplot2}). As such, most of these plotting tools can be customized by appending syntax from the \pkg{ggplot2} package, as will be demonstrated in example code throughout this paper. Moreover, some of these plotting tools have interactive capabilities that were developed with the use of the \pkg{plotly} package (\citealt{plotly}) (see Section \ref{interaction}).

@@

\begin{Code}
R> plotPath(pathTN, sbGeneal, "devYear")
\end{Code}

@@

\begin{Code}
R> pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal, "devYear")
R> plotPath(pathBZ, sbGeneal, "devYear", fontFace = 2) +
+    ggplot2::xlab("Development Year")
\end{Code}

@@

\begin{Code}
R> sbFilt <- sbGeneal[complete.cases(sbGeneal[1:3]),]
R> sbFiltIG <- dfToIG(sbFilt)
R> pathCL <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "yield")
R> plotPath(pathCL, sbFilt, "yield") + ggplot2::xlab("Yield")
R> pathCL2 <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "devYear")
R> plotPath(pathCL2, sbFilt, "devYear", "yield") +
+    ggplot2::xlab("Development Year") + ggplot2::ylab("Yield")
\end{Code}

@@

\begin{Code}
R> plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 3, pathEdgeCol =
+    "red", nodeSize = 2.5, pathNodeSize = 4) +
+    ggplot2::theme(axis.text = ggplot2::element_text(size = 12),
+    axis.title = ggplot2::element_text(size = 12))
\end{Code}

@@

\begin{Code}
R> plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 6, pathEdgeCol =
+    "seagreen2", nodeSize = 1, pathNodeSize = 3) +
+    ggplot2::xlab("Development Year")
\end{Code}

@@

We would like to emphasize that the \code{plotPathOnAll()} method can be created with one or two quantitative variables that the user hard-codes. We demonstrate this using the filtered data frame and igraph objects (\code{sbFilt} and \code{sbFiltIG}) that we previously created to assure that there were no NA values in the quantitative variable ``yield". Then, we can examine the path across the remaining genealogical structure and how the quantitative variable ``yield" changes along the parent-child relationships of that path. Likewise, we can plot how the quantitative variables ``devYear" and ``yield" both change along the parent-child relationships of that path. The output of these two calls to the \code{plotPathOnAll()} function can be viewed in Figure \ref{fig:plotPathOnAllQV2}.

\begin{Code}
R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", bin = 3, pathEdgeCol =
+    "purple") + ggplot2::xlab("Yield")
R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear",
+    pathEdgeCol = "orange") + ggplot2::xlab("Yield") +
+    ggplot2::ylab("Development Year")
\end{Code}

@@

\begin{Code}
R> plotAncDes("Lee", sbGeneal, mAnc = 6, mDes = 6, vCol = "blue")
\end{Code}

@@

\begin{Code}
>R varieties <- c("Brim", "Bedford", "Calland", "Dillon", "Hood", "Narow",
+    "Pella", "Tokyo", "Young", "Zane")
>R plotDegMatrix(varieties, sbIG, sbGeneal) +
+    ggplot2::scale_fill_continuous(low = "white", high = "darkgreen") +
+    ggplot2::theme(legend.title = ggplot2::element_text(size = 15),
+    legend.text = ggplot2::element_text(size = 15)) +
+    ggplot2::labs(x = "Variety", y = "Variety")
\end{Code}

@@

\begin{CodeChunk}
\begin{CodeInput}
>R library("dplyr")
>R indVec <- getNodes(statGeneal)
>R indVec <- indVec[which(indVec != "", )]
>R dFunc <- function(var) nrow(getDescendants(var, statGeneal, gen = 100))
>R numDesc <- sapply(indVec, dFunc)
>R table(numDesc)
\end{CodeInput}
\begin{CodeOutput}
numDesc
   0    1    2    3    4    5    6    7    8    9   10   11   12   13   14
6251  322  145   88   58   36   31   22   23   14   17   13   14   10    9
  15   16   17   18   19   20   21   22   23   24   25   26   27   29   30
   6    4    3    2    5    7    4    3    3    2    2    6    1    1    3
  34   37   38   40   41   44   45   48   49   60   61   75   77   84  159
   2    1    1    1    1    1    1    1    2    1    1    1    1    1    1
\end{CodeOutput}
\begin{CodeInput}
R> which(numDesc == 159)
\end{CodeInput}
\begin{CodeOutput}
David Cox
     1980
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> statIG <- dfToIG(statGeneal)
R> pathCB <- getPath("David Cox", "Petra Buzkova", statIG, statGeneal,
+    "gradYear", isDirected = FALSE)
R> plotPath(pathCB, statGeneal, "gradYear", fontFace = 4) +
+    ggplot2::xlab("Graduation Year") + ggplot2::theme(axis.text =
+    ggplot2::element_text(size = 10), axis.title =
+    ggplot2::element_text(size = 10)) +
+    ggplot2::scale_x_continuous(expand = c(.1, .2))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200) +
+    ggplot2::theme(axis.text = ggplot2::element_text(size = 8),
+    axis.title = ggplot2::element_text(size = 8)) +
+    ggplot2::scale_x_continuous(expand = c(.1, .2)) +
+    ggplot2::xlab("Graduation Year")
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200,
+    nodeSize = .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol =
+    "lightgray") + ggplot2::theme(axis.text = ggplot2::element_text(size = +    8), axis.title = ggplot2::element_text(size = 8)) +
+    ggplot2::scale_x_continuous(expand = c(.1, .2)) +
+    ggplot2::xlab("Graduation Year")
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200,
+    nodeSize = .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol =
+    "lightgray", animate = TRUE)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{Code}
R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", pathEdgeCol = "orange",
+    animate = TRUE)
R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear",
+    pathEdgeCol = "orange", animate = TRUE)
\end{Code}

>>>>

Additionally, the \pkg{ggenealogy} package allows for users to plot genealogical data in multiple ways. Users can obtain and plot the shortest path between two nodes of interest constrained on a variable of interest (see Section \ref{psp}), superimpose a shortest path between two nodes of interest with the entire genealogical structure constrained on a variable of interest (see Section \ref{sspt}), plot the ancestors and descendants of a node of interest constrained on generation count (see Section \ref{remedy}), and plot distance matrices based on a variable of interest (see Section \ref{pdm}). Most of these plotting tools were developed by building upon the \pkg{ggplot2} package (\citealt{ggplot2}). As such, most of these plotting tools can be customized by appending syntax from the \pkg{ggplot2} package, as will be demonstrated in example code throughout this paper. Moreover, some of these plotting tools have interactive capabilities that were developed with the use of the \pkg{plotly} package (\citealt{plotly}) (see Section \ref{interaction}).

@@

\begin{CodeInput}
R> plotPath(pathTN, sbGeneal, "devYear")
\end{CodeInput}

@@

\begin{CodeInput}
R> pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal, "devYear")
R> plotPath(pathBZ, sbGeneal, "devYear", fontFace = 2) +
+    ggplot2::xlab("Development Year")
\end{CodeInput}

@@

\begin{CodeInput}
R> sbFilt <- sbGeneal[complete.cases(sbGeneal[1:3]), ]
R> sbFiltIG <- dfToIG(sbFilt)
R> pathCL <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "yield")
R> plotPath(pathCL, sbFilt, "yield") + ggplot2::xlab("Yield")
R> pathCL2 <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "devYear")
R> plotPath(pathCL2, sbFilt, "devYear", "yield") +
+    ggplot2::xlab("Development Year") + ggplot2::ylab("Yield")
\end{CodeInput}

@@

\begin{CodeInput}
R> plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 3, pathEdgeCol =
+    "red", nodeSize = 2.5, pathNodeSize = 4) +
+    ggplot2::theme(axis.text = ggplot2::element_text(size = 12),
+    axis.title = ggplot2::element_text(size = 12))
\end{CodeInput}

@@

\begin{CodeInput}
R> plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 6, pathEdgeCol =
+    "seagreen2", nodeSize = 1, pathNodeSize = 3) +
+    ggplot2::xlab("Development Year")
\end{CodeInput}

@@

We would like to emphasize that the \code{plotPathOnAll()} method can be created with one or two quantitative variables that the user hard-codes. We demonstrate this using the filtered data frame and igraph objects (\code{sbFilt} and \code{sbFiltIG}) to assure that there were no NA values in the quantitative variable ``yield". Then, we can examine the path across the remaining genealogical structure and how the quantitative variable ``yield" changes along the parent-child relationships of that path. Likewise, we can plot how the quantitative variables ``devYear" and ``yield" both change along the parent-child relationships of that path. The output of these two calls to the \code{plotPathOnAll()} function can be viewed in Figure \ref{fig:plotPathOnAllQV2}.

\begin{CodeInput}
R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", bin = 3, pathEdgeCol =
+    "purple") + ggplot2::xlab("Yield")
R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear",
+    pathEdgeCol = "orange") + ggplot2::xlab("Yield") +
+    ggplot2::ylab("Development Year")
\end{CodeInput}

@@

\begin{CodeInput}
R> plotAncDes("Lee", sbGeneal, mAnc = 6, mDes = 6, vCol = "blue")
\end{CodeInput}

@@

\begin{CodeInput}
R> varieties <- c("Brim", "Bedford", "Calland", "Dillon", "Hood", "Narow",
+    "Pella", "Tokyo", "Young", "Zane")
R> plotDegMatrix(varieties, sbIG, sbGeneal) +
+    ggplot2::scale_fill_continuous(low = "white", high = "darkgreen") +
+    ggplot2::theme(legend.title = ggplot2::element_text(size = 15),
+    legend.text = ggplot2::element_text(size = 15)) +
+    ggplot2::labs(x = "Variety", y = "Variety")
\end{CodeInput}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> library("dplyr")
R> indVec <- getNodes(statGeneal)
R> indVec <- indVec[which(indVec != "", )]
R> dFunc <- function(var) nrow(getDescendants(var, statGeneal, gen = 100))
R> numDesc <- sapply(indVec, dFunc)
R> table(numDesc)
\end{CodeInput}
\begin{CodeOutput}
numDesc
   0    1    2    3    4    5    6    7    8    9   10   11   12   13   14
6251  322  145   88   58   36   31   22   23   14   17   13   14   10    9
  15   16   17   18   19   20   21   22   23   24   25   26   27   29   30
   6    4    3    2    5    7    4    3    3    2    2    6    1    1    3
  34   37   38   40   41   44   45   48   49   60   61   75   77   84  159
   2    1    1    1    1    1    1    1    2    1    1    1    1    1    1
\end{CodeOutput}
\begin{CodeInput}
R> which(numDesc == 159)
\end{CodeInput}
\begin{CodeOutput}
David Cox
     1980
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> statIG <- dfToIG(statGeneal)
R> pathCB <- getPath("David Cox", "Petra Buzkova", statIG, statGeneal,
+    "gradYear", isDirected = FALSE)
R> plotPath(pathCB, statGeneal, "gradYear", fontFace = 4) +
+    ggplot2::xlab("Graduation Year") + ggplot2::theme(axis.text =
+    ggplot2::element_text(size = 10), axis.title =
+    ggplot2::element_text(size = 10)) +
+    ggplot2::scale_x_continuous(expand = c(.1, .2))
\end{CodeInput}

@@

\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200) +
+    ggplot2::theme(axis.text = ggplot2::element_text(size = 8),
+    axis.title = ggplot2::element_text(size = 8)) +
+    ggplot2::scale_x_continuous(expand = c(.1, .2)) +
+    ggplot2::xlab("Graduation Year")
\end{CodeInput}

@@

\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200,
+    nodeSize = .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol =
+    "lightgray") + ggplot2::theme(axis.text = ggplot2::element_text(size =
+    8), axis.title = ggplot2::element_text(size = 8)) +
+    ggplot2::scale_x_continuous(expand = c(.1, .2)) +
+    ggplot2::xlab("Graduation Year")
\end{CodeInput}

@@

\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200,
+    nodeSize = .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol =
+    "lightgray", animate = TRUE)
\end{CodeInput}

@@

\begin{CodeInput}
R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", pathEdgeCol = "orange",
+    animate = TRUE)
R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear",
+    pathEdgeCol = "orange", animate = TRUE)
\end{CodeInput}
